---
title: "Aedes Vignette"
author: "Michael Johansson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aedes Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(predx)
```

This package contains tools for working with prediction data. Much of the functionality is still in development, but forecast validation has already been implemented for the Aedes Forecasting Challenge.

## Import Aedes Forecasting Challenge forecast

First, the forecaster should create their forecast in the CSV format specified for the Aedes Forecast Challenge by the CDC Epidemic Prediction Initiative (provided as a template at [predict.cdc.gov](https://predict.cdc.gov)). Forecasts matching this template can be imported using the `import_aedes_csv` function. This function imports the CSV and converts it to a special, embedded "`predx_df`" data frame (a `tbl_df`, "tibble" object) with a row for every individual prediction (defined by target, location, and prediction type), a column `predx_class` that defines the class of each prediction, and a column `predx` that is a list of predx objects. In the process of importation, all predictions are validated (e.g.probabilities can't be negative). If conversion fails, (hopefully useful) errors messages are returned in the `predx` column.

This function ingests a CSV file, converts the columns to those needed to create a `predx` object, and creates a `predx_df` object (throwing error messages if conversion fails).

If the forecast will be stored in a different format, it may be helpful to include additional variables in the `predx_df` (e.g. submission date, month forecasted). These can be included as additional columns by providing a list to the optional argument `add_vars` as shown below. 

```{r}
fcast <- import_aedes_csv('aedes_null_forecast.csv', 
  add_vars=list(due_date='2019-03-31', month='April', team='null'))
class(fcast)
fcast
```

## Verify expected predictions are included

The `verify_expected` function can be used to check that expected targets are included in the `predx_df`. For the Aedes challenge, this includes Binary forecasts for two targets and 95 counties. These are specified in a list returned from `aedes_expected` below (note that it is a 2-level list to allow for additional combination).
```
list(
  list(
    target = c("Ae. aegypti", "Ae. albopictus"),
    location =
      c("California-Alameda", "California-Butte", "California-Colusa",
      "California-Contra Costa", "California-Fresno", "California-Glenn",
      "California-Imperial", "California-Inyo", "California-Kern",
      "California-Kings", "California-Lake", "California-Los Angeles",
      "California-Madera", "California-Marin", "California-Merced",
      "California-Mono", "California-Monterey", "California-Napa",
      "California-Orange", "California-Placer", "California-Riverside",
      "California-Sacramento", "California-San Benito", "California-San Bernardino",
      "California-San Diego", "California-San Francisco", "California-San Joaquin",
      "California-San Luis Obispo", "California-San Mateo", "California-Santa Barbara",
      "California-Santa Clara", "California-Santa Cruz", "California-Shasta",
      "California-Solano", "California-Sonoma", "California-Stanislaus",
      "California-Sutter", "California-Tulare", "California-Ventura",
      "California-Yolo", "California-Yuba", "Connecticut-Fairfield",
      "Connecticut-New Haven", "Florida-Calhoun", "Florida-Collier",
      "Florida-Escambia", "Florida-Gadsden", "Florida-Hillsborough",
      "Florida-Holmes", "Florida-Jackson", "Florida-Jefferson", "Florida-Lee",
      "Florida-Liberty", "Florida-Madison", "Florida-Manatee", "Florida-Martin",
      "Florida-Miami-Dade", "Florida-Okaloosa", "Florida-Osceola",
      "Florida-Pasco", "Florida-Pinellas", "Florida-Polk", "Florida-Santa Rosa",
      "Florida-St. Johns", "Florida-Taylor", "Florida-Wakulla", "Florida-Walton",
      "Florida-Washington", "New Jersey-Cumberland", "New Jersey-Essex",
      "New Jersey-Mercer", "New Jersey-Monmouth", "New Jersey-Morris",
      "New Jersey-Salem", "New Jersey-Sussex", "New Jersey-Warren",
      "New York-Bronx", "New York-Kings", "New York-Nassau", "New York-New York",
      "New York-Queens", "New York-Richmond", "New York-Rockland",
      "New York-Westchester", "North Carolina-Forsyth", "North Carolina-New Hanover",
      "North Carolina-Pitt", "North Carolina-Transylvania", "North Carolina-Wake",
      "Texas-Cameron", "Texas-Hidalgo", "Texas-Tarrant", "Wisconsin-Dane",
      "Wisconsin-Milwaukee", "Wisconsin-Waukesha"),
    predx_class = "Binary"
  )
)
```

This specific expected list, `aedes_expected`, is included in the `predx` package. Other verification lists can be made in this format and all can be used with the function `verify_expected` to validate that all expected predictions are included. The function prints missing and additional predictions, but those can also be returned as a data frame by including the argument `return_df = TRUE`.

```{r}
verify_expected(fcast, aedes_expected())
```

## Export as `predx` JSON (reduced size) or `predx` CSV

To save space and facilitate sharing, transfer, and storage of `predx_df` objects, they can be exported as JSON objects using the function `export_json`. A file name can be supplied as an argument to store this as a file instead of returning an object as shown for the first two predictions in `fcast` below.

```{r}
export_json(fcast[1:2, ])
json_tempfile = tempfile()
export_json(fcast, filename = json_tempfile, overwrite = T)
```

Alternatively, `predx_df` objects can be exported as predx CSV files. This CSV formats differs slightly from the Aedes Forecasting Challenge template because it may contains additional columns (e.g. predx_class) and it does not contain the required column 'value'. The similar function, `export_aedes_csv` adds the value column so that the CSV can be uploaded to the Epidemic Prediction Initiative website (the extra columns are kept, but not interpreted by the website). 

It may be helpful to generate forecasts in the `predx_df` format. Any forecasts generated in that format can be exported to the Aedes Forecasting Challenge submission format using `export_aedes_csv` (provided they include the required columns "type" and "unit" and have been verified using `verify_expected`).

```{r}
export_csv(fcast[1:2, ])
export_aedes_csv(fcast[1:2, ])
csv_tempfile <- tempfile()
export_aedes_csv(fcast, filename = csv_tempfile, overwrite = T)
```

## Import `predx` JSON or `predx` CSV

In addition to importing CSV files in the Aedes Forecasting Challenge format (as described at the beginning of this vignette), `predx` can be used to import `predx` JSON or `predx` CSV files.

```{r}
fcast <- import_json(json_tempfile)
head(fcast)
fcast_csv <- import_csv(csv_tempfile)
head(fcast_csv)
```

