---
title: "FluSight Vignette"
author: "Michael Johansson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flusight vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Import FluSight forecast

The `import_flusight_csv` function uses several helper functions to convert FluSight-specific formatted CSV forecasts. The file name should contain the MMWR epidemiological week assigned to the forecast (e.g. "EW42") followed by a dash ("-"), the team name (e.g. "Hist_Avg") followed by a dash ("-"), and the submission date as an ISO data (YYYY-MM-DD, e.g. "2018-10-29").

This import process creates a special, embedded "`predx_df`" data frame (a `tbl_df`, "tibble" object) with a row for every individual prediction (defined by target, location, and prediction type), a column `predx_class` that defines the class of each prediction, and a column `predx` that is a list of predx objects. In the process of importation, all predictions are validated (e.g. binned predictions must sum to 1.0) or (hopefully useful) errors messages are returned in the `predx` column.

For this example, we use a submission with only national-level forecasts.

```{r}
library(predx)
fcast <- import_flusight_csv('EW42-Hist-Avg-2018-10-29-National.csv')
class(fcast)
fcast
```

## Verify expected predictions are included

The `verify_expected` function can be used to check that expected targets are included in the `predx_df`. For example, the national-level FluSight forecasts include forecasts for seven targets and 11 locations. The percentage forecasts (e.g. "Season peak percentage", "1 wk ahead") are classified as `BinLwr` predictions and the week targets ("Season onset", "Season peak week") are classifed as `BinCat` predictions. This set can be specified in a list with one list for each combination of required characteristics. For example, the target "Season peak percentage", should have a prediction for each location and each predx bin as specified in the first element of `flusight_ilinet_expected()`, below.

```
list(
  list(
    target = c("Season peak percentage", "1 wk ahead", "2 wk ahead", 
      "3 wk ahead", "4 wk ahead"),
    location = c("HHS Region 1", "HHS Region 10", "HHS Region 2", "HHS Region 3",
        "HHS Region 4", "HHS Region 5", "HHS Region 6", "HHS Region 7",
        "HHS Region 8", "HHS Region 9", "US National"),
    predx_class = "BinLwr",
    lwr = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.1, 1.2, 
      1.3, 1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2, 2.1, 2.2, 2.3, 2.4, 2.5, 
      2.6, 2.7, 2.8, 2.9, 3, 3.1, 3.2, 3.3, 3.4, 3.5, 3.6, 3.7, 3.8, 
      3.9, 4, 4.1, 4.2, 4.3, 4.4, 4.5, 4.6, 4.7, 4.8, 4.9, 5, 5.1, 
      5.2, 5.3, 5.4, 5.5, 5.6, 5.7, 5.8, 5.9, 6, 6.1, 6.2, 6.3, 6.4, 
      6.5, 6.6, 6.7, 6.8, 6.9, 7, 7.1, 7.2, 7.3, 7.4, 7.5, 7.6, 7.7, 
      7.8, 7.9, 8, 8.1, 8.2, 8.3, 8.4, 8.5, 8.6, 8.7, 8.8, 8.9, 9, 
      9.1, 9.2, 9.3, 9.4, 9.5, 9.6, 9.7, 9.8, 9.9, 10, 10.1, 10.2, 
      10.3, 10.4, 10.5, 10.6, 10.7, 10.8, 10.9, 11, 11.1, 11.2, 11.3, 
      11.4, 11.5, 11.6, 11.7, 11.8, 11.9, 12, 12.1, 12.2, 12.3, 12.4, 
      12.5, 12.6, 12.7, 12.8, 12.9, 13)
  ),
  list(
    target = "Season peak week",
    location = c("HHS Region 1", "HHS Region 10", "HHS Region 2", "HHS Region 3",
        "HHS Region 4", "HHS Region 5", "HHS Region 6", "HHS Region 7",
        "HHS Region 8", "HHS Region 9", "US National"),
    predx_class = "BinCat",
    cat = c("40", "41", "42", "43", "44", "45", "46", "47", "48", "49", 
      "50", "51", "52", "1", "2", "3", "4", "5", "6", "7", "8", "9", 
      "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20")
    ),
  list(
    target = "Season onset",
    location = c("HHS Region 1", "HHS Region 10", "HHS Region 2", "HHS Region 3",
        "HHS Region 4", "HHS Region 5", "HHS Region 6", "HHS Region 7",
        "HHS Region 8", "HHS Region 9", "US National"),
    predx_class = "BinCat",
    cat = c("40", "41", "42", "43", "44", "45", "46", "47", "48", "49", 
      "50", "51", "52", "1", "2", "3", "4", "5", "6", "7", "8", "9", 
      "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", 
      "none")
    ),
  list(
    target = c("Season peak percentage", "1 wk ahead", "2 wk ahead", 
      "3 wk ahead", "4 wk ahead", "Season onset", "Season peak week"),
    location = c("HHS Region 1", "HHS Region 10", "HHS Region 2", "HHS Region 3",
        "HHS Region 4", "HHS Region 5", "HHS Region 6", "HHS Region 7",
        "HHS Region 8", "HHS Region 9", "US National"),
    predx_class = "Point"
    )
```

This specific expected list, available with `flusight_ilinet_expected`, is included in the predx package along with `flusight_state_ilinet_expected` (for state ILINet forecasts) and `flusight_hospitalization_expected` (for hospitalization forecasts). Other verification lists can be made in this format and all can be used with the function `verify_expected` to validate that all expected predictions are included. The function prints missing and additional predictions, but those can also be returned as a data frame by including the argument `return_df = TRUE`.

For this example, we limit the expected forecasts to the national level.

```{r}
national_expected <- list(
  list(
    target = c("Season peak percentage", "1 wk ahead", "2 wk ahead", 
      "3 wk ahead", "4 wk ahead"),
    location = "US National",
    predx_class = "BinLwr",
    lwr = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.1, 1.2, 
      1.3, 1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2, 2.1, 2.2, 2.3, 2.4, 2.5, 
      2.6, 2.7, 2.8, 2.9, 3, 3.1, 3.2, 3.3, 3.4, 3.5, 3.6, 3.7, 3.8, 
      3.9, 4, 4.1, 4.2, 4.3, 4.4, 4.5, 4.6, 4.7, 4.8, 4.9, 5, 5.1, 
      5.2, 5.3, 5.4, 5.5, 5.6, 5.7, 5.8, 5.9, 6, 6.1, 6.2, 6.3, 6.4, 
      6.5, 6.6, 6.7, 6.8, 6.9, 7, 7.1, 7.2, 7.3, 7.4, 7.5, 7.6, 7.7, 
      7.8, 7.9, 8, 8.1, 8.2, 8.3, 8.4, 8.5, 8.6, 8.7, 8.8, 8.9, 9, 
      9.1, 9.2, 9.3, 9.4, 9.5, 9.6, 9.7, 9.8, 9.9, 10, 10.1, 10.2, 
      10.3, 10.4, 10.5, 10.6, 10.7, 10.8, 10.9, 11, 11.1, 11.2, 11.3, 
      11.4, 11.5, 11.6, 11.7, 11.8, 11.9, 12, 12.1, 12.2, 12.3, 12.4, 
      12.5, 12.6, 12.7, 12.8, 12.9, 13)
  ),
  list(
    target = "Season peak week",
    location = "US National",
    predx_class = "BinCat",
    cat = c("40", "41", "42", "43", "44", "45", "46", "47", "48", "49", 
      "50", "51", "52", "1", "2", "3", "4", "5", "6", "7", "8", "9", 
      "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20")
    ),
  list(
    target = "Season onset",
    location = "US National",
    predx_class = "BinCat",
    cat = c("40", "41", "42", "43", "44", "45", "46", "47", "48", "49", 
      "50", "51", "52", "1", "2", "3", "4", "5", "6", "7", "8", "9", 
      "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", 
      "none")
    ),
  list(
    target = c("Season peak percentage", "1 wk ahead", "2 wk ahead", 
      "3 wk ahead", "4 wk ahead", "Season onset", "Season peak week"),
    location = "US National",
    predx_class = "Point"
    )
)

verify_expected(fcast, national_expected)
```

## Export as predx JSON (reduced size), predx CSV, or FluSight CSV

To save space and facilitate sharing, transfer, and storage of `predx_df` objects, they can be exported as JSON objects using the function `export_json`. A file name can be supplied as an argument to store this as a file instead of returning an object as shown for the peak week predictions in `fcast` below.

```{r}
peak_week_json <- export_json(fcast[13:14, ])
peak_week_json
jsonlite::prettify(peak_week_json)
json_tempfile = tempfile()
export_json(fcast, filename = json_tempfile, overwrite = T)
```

Alternatively, `predx_df` objects can be exported as predx CSV files or FluSight-formatted CSV files (those used for submissions on the FluSight website). These CSV formats differ in several ways: 

1. The `predx` CSV contains additional columns for MMWR epidemic week, team, and submission date. In the FluSight CSV these are included in the file name only. 
2. The `predx` CSV includes point predictions as individual rows whereas the FluSight CSV includes them in an additional column. 
3. The FluSight CSV will allow upload on the EPI webpage.

These differences are the results of making `predx` JSON and CSV formats more generic than the specific FluSight implementation.

```{r}
csv_tempfile <- tempfile()
export_csv(fcast, filename = csv_tempfile, overwrite = T)
export_flusight_csv(fcast[1:2, ])
```

## Import predx JSON or predx CSV

In addition to importing CSV files in the FluSight format (as described at the beginning of this vignette), `predx` can be used to import `predx` JSON or `predx` CSV files.

```{r}
fcast <- import_json(json_tempfile)
head(fcast)
fcast_csv <- import_csv(csv_tempfile)
head(fcast_csv)
```

## Use with the FluSight package

The `FluSight` package [https://github.com/jarad/FluSight]() includes numerous functions to work with forecasts, such as scoring and visualization. `predx` FluSight forecast can be converted to the format used by the `FluSight` package using `to_flusight_pkg_format`.

```{r}
fcast_flusight <- to_flusight_pkg_format(fcast)
head(fcast_flusight)
```

This can then be used for scoring. Note that in this example, `fcast` only contains national level forecasts.

```{r}
library(FluSight)
truth_1819 <- FluSight::create_truth(year = 2018)
truth_1819 <- dplyr::filter(truth_1819, location == 'US National')
FluSight::score_entry(fcast_flusight, truth_1819)
```



